---
title: "parse_kinetic_data"
---
Import required packages
```{r}
library(plyr,include.only = c('mapvalues'))
library(dplyr)

library(readxl)
library(sybil)
library(stringr)
```

First, we load the Heckman et al (2020) data
```{r}
#k_app data for growth on glucose
k_app_data=read_xlsx("../Heckman_et_al_2020/Heckman_et_al_2020_SI_data.xlsx",
                     sheet='Dataset_S1A_protein_ab')
#k_app_max data (in-vivo proxy for kcat)
k_app_max_data=read_xlsx("../Heckman_et_al_2020/Heckman_et_al_2020_SI_data.xlsx",
                     sheet='Dataset_S1C_turnover_n')

```

Next, we load the enzymatic reactions in the model
```{r}
enzymatic_reactions=read.csv('../../../Model/model_tables/enzymatic_reactions.csv',header = F) %>% .$V1
enzymatic_reactions %>% head(10)
```
First, let's deal with k_app (glucose). We remove NAs and only retain wild-type. Further we remove the _b suffix for backward reactions, as that info is already contained in the "direction" column
```{r}
wt_k_apps=
  k_app_data %>% 
  filter(sample_id %in% c('WT1','WT2')) %>% 
  filter(kapp_per_s!= 'NA') %>% 
  mutate(react_id=str_replace(react_id,pattern='_b$',replacement='')) %>% 
  filter(react_id %in% enzymatic_reactions) %>% 
  mutate(kapp_per_s=as.numeric(kapp_per_s))
```

Next, we summarise data. We take each k_app to be the median of of all k_app available
```{r}
model_k_apps =
  wt_k_apps %>% 
  group_by(react_id,direction) %>% 
  summarise(median_kapp_per_s=median(kapp_per_s)) %>% 
  mutate(par_type='k_app',unit='s-1',condition='Glucose',notes='median of calculated k_app') %>% 
  rename(value=median_kapp_per_s) 

model_k_apps  
```
How many kapps do we have?
```{r}
print(paste('Number of reactions with kapp in the model:',model_k_apps$react_id %>% unique() %>% length()))
```
Now, we parse k_app_max
```{r}
model_k_app_max=
  k_app_max_data %>% 
  mutate(is_backward=grepl(pattern='_b$',x=react_id)) %>% 
  mutate(direction=mapvalues(x=is_backward,from=c(T,F),to=c('bwd','fwd')))%>% 
  mutate(react_id=str_replace(react_id,pattern='_b$',replacement='')) %>% 
  select(react_id,direction,kappmax_KO_ALE_per_pp_per_s_ensemble_model) %>% 
  rename(value=kappmax_KO_ALE_per_pp_per_s_ensemble_model) %>% 
  mutate(par_type='k_appmax',unit='s-1',condition='NA',notes='kappmax_KO_ALE_per_pp_per_s_ensemble_model')
  
```

Merge the two
```{r}
heckmann_parsed=rbind(model_k_apps,model_k_app_max)
heckmann_parsed %>% 
  rename(reaction_id=react_id) %>% 
  write.csv(file='parsed/heckman_et_al_2020_parsed_kapp_kappmax.csv',row.names = F)
```


