---
title: "R Notebook"
---
In this notebook, we parse turnover numbers and MW into tables which will be used to set-up sMoment formulation of the model

```{r}
#Load required packages
library(plyr,include.only = 'mapvalues')
library(dplyr)
library(tidyr)
library(stringr)
library(reshape2)
USE_ONLY_PRIMARY=T
if (USE_ONLY_PRIMARY){
   admissible_edge_types=c('catalysis')
   
}else{
  admissible_edge_types=c('catalysis','secondary_catalysis')
  }


```
Read turnover number and MW data, and enzymatic reactions in model

```{r}
kcat_data=read.csv('parsed/heckman_et_al_2020_parsed_kapp_kappmax.csv') %>% 
  mutate(direction=gsub(pattern='bwd',replacement='bw',x=direction)) %>% 
  mutate(direction=gsub(pattern='fwd',replacement='fw',x=direction)) 
kcat_data[is.na(kcat_data)]="NA"
  
kcat_manual_curation=read.csv('kcat_manual_curation.csv')
mw_data=read.csv('../../../Model/model_tables/reaction_enzymes_subunits.csv')
enzymatic_reactions=read.csv('../../../Model/model_tables/enzymatic_reactions.csv',header = F) %>% .$V1
iml1515_transporters=
  read.csv('../Heckman_et_al_2020/iML1515_list_of_transporters.csv') %>% 
  mutate(reaction_id=str_replace(bigg_id,pattern='_[b,f]$',replacement=''))
paxDB_estimated_abundance=
  read.csv('../../enzyme_abundance_estimation/schmidt_2016_mass_abundance_mapped_NNLS.csv') %>% 
  filter(enzyme %in% mw_data$enzyme) %>%
  select(c('enzyme','paxDB_g_gDW'))
mw_data$paxDB_abundance=mapvalues(mw_data$enzyme,from=paxDB_estimated_abundance$enzyme,to=paxDB_estimated_abundance$paxDB_g_gDW) %>% as.numeric()

```

Now, following the sMoment formulation, we seek to choose an individual enzyme per reaction. To maximize physiological relevance, we only consider primary catalysis.

Next, following sMOMENT, we pick the one with least MW, as it will be the one chosen in any proteome-limited solution to the model

```{r}

reaction_enzyme_unique=
  mw_data %>% 
  mutate(paxDB_abundance=replace_na(paxDB_abundance,0)) %>% 
  filter(edge_type %in% admissible_edge_types) %>% 
  group_by(reaction_id) %>% 
  slice_max(paxDB_abundance)
reaction_enzyme_unique %>% head()
```
### Assemble Glucose-specific and general EC-Model Tables. 
For the glucose-specific model we choose:
- A kapp, if available in the dataset
- A kapp_max if the kapp for the reaction is unavailable

For the general EC-model, we always choose kapp_maxs

```{r}
default_transporter_kapp=65
default_kappmax_saturation=1


turnover_table=data.frame()
for (model_type in c('glucose_batch','general')){
  for(reaction in enzymatic_reactions){
    #If we have a kapp, use this
    if(model_type=='glucose_batch' & (reaction %in% (kcat_data %>% filter(par_type=='k_app') %>% .$reaction_id))){
       turnover_table =rbind(turnover_table,
                           kcat_data %>% 
                             filter(par_type=='k_app') %>%
                             filter(reaction_id==reaction) %>% 
                             mutate(saturation=1,model_type=model_type))
    }
    
    #If we have a kappmax, use this
    else 
    if(reaction %in% (kcat_data %>% filter(par_type=='k_appmax') %>% .$reaction_id)){
       turnover_table =rbind(turnover_table,
                                   kcat_data %>% filter(par_type=='k_appmax') %>% 
                                     filter(reaction_id==reaction) %>% 
                                     mutate(saturation=default_kappmax_saturation,model_type=model_type))
    }
    else
      if(reaction %in% iml1515_transporters$reaction_id){
        temp_transporter_df=
          data.frame(      reaction_id=c(reaction,reaction),
                           direction=c('fw','bw'),
                           value=c(default_transporter_kapp,default_transporter_kapp),
                           par_type=c(default_transporter_kapp,default_transporter_kapp),
                           unit=c('s-1','s-1'),
                           condition=c('NA','NA'),
                           notes=c('same as Heckman et al. 2018','same as Heckman et al. 2018'),
                           saturation=c(default_kappmax_saturation,default_kappmax_saturation),
                           model_type=c(model_type,model_type)
                           )
       turnover_table=rbind(turnover_table,temp_transporter_df)
      }
    #If we have a manual curation, let's
    if(reaction %in% (kcat_manual_curation %>% .$reaction_id)){
       turnover_table =rbind(turnover_table,
                             kcat_manual_curation %>% 
                            select(!enzyme) %>% 
                             filter(reaction_id==reaction) %>% 
                             mutate(saturation=default_kappmax_saturation,model_type=model_type)
                             )
    }
    }
}
```

Now, we add info about mapped enzyme, MW, and number of subunits
```{r}
EC_table=
  turnover_table %>% 
  rename(reaction_id=reaction_id) %>% 
  mutate(directional_reaction_id=paste(reaction_id,direction,sep='_')) %>% 
  mutate(enzyme=mapvalues(reaction_id,from=reaction_enzyme_unique$reaction_id,to=reaction_enzyme_unique$enzyme,warn_missing=F),
         MW=mapvalues(reaction_id,from=reaction_enzyme_unique$reaction_id,to=reaction_enzyme_unique$mw,warn_missing=F) %>% 
           as.numeric(),
         num_pp_subunits=mapvalues(reaction_id,from=reaction_enzyme_unique$reaction_id,to=reaction_enzyme_unique$num_subunits,warn_missing=F) %>% 
           as.numeric()
         ) %>% 
  filter(!is.na(MW)) %>% 
  mutate(kcat_times_pp_num=value*num_pp_subunits)


EC_table$EC_kcat_value=EC_table$kcat_times_pp_num
#EC_table[(EC_table$condition =='manual_curation') %>% replace(is.na(.),F),'EC_kcat_value']=EC_table[(EC_table$condition =='manual_curation')%>% replace(is.na(.),F),'value']
#For complexes with more than 20 subunits, do not multiply the kcat times the number of subunits, as this would result in unrealistically fast catalysis
# large_enzyme_complexes=EC_table %>% filter(num_pp_subunits>20) %>% .$enzyme %>% unique()
# 
# EC_table$EC_kcat_value=EC_table$kcat_times_pp_num
# EC_table[EC_table$enzyme %in% large_enzyme_complexes,'EC_kcat_value']=EC_table[EC_table$enzyme %in% large_enzyme_complexes,'value']
```
Finally, we compute the cost in enzyme_mass per flux

```{r}
EC_table %>% 
  mutate(ec_cost=MW/(saturation*EC_kcat_value*3600))->
  EC_table


EC_table %>% head()
```
Finally, we compute a default proteomic bound based on batch growth on glucose
```{r}
proteomic_data_glucose=read.csv('../Schmidt_et_al_2016/model_mapped_abundances/schmidt_2016_mass_abundance_mapped_avg.csv') %>% select(c('enzyme','glucose_g_gDW'))

measured_model_total_abundance=
  proteomic_data_glucose %>% 
  group_by(enzyme) %>% 
  slice(1) %>% 
  ungroup() %>% 
  .$glucose_g_gDW %>% 
  sum(na.rm = T)

EC_table=
  EC_table%>% 
  mutate(enzyme_pool_bound_g_gDw=measured_model_total_abundance) 

  
EC_table_no_manual_curation=
  EC_table %>% 
  filter(condition!="manual_curation")
 
if (USE_ONLY_PRIMARY){
EC_table_no_manual_curation %>% write.csv('parsed/EC_table.csv',row.names = F)
}else{
  EC_table_no_manual_curation %>% write.csv('parsed/EC_table_w_secondary_edges.csv',row.names = F)
}


```

Also, save a table with manual curation values over kappmax, whenever both exists
```{r}
EC_table_manual_curation=
  EC_table %>% 
  group_by(reaction_id,enzyme,direction) %>% 
  filter(if ('manual_curation' %in% condition) condition=='manual_curation' else T) %>%
  ungroup()
if (USE_ONLY_PRIMARY){
EC_table_manual_curation %>% write.csv('parsed/EC_table_w_manual_curation.csv',row.names = F)
}else{
EC_table_manual_curation %>% write.csv('parsed/EC_table_w_manual_curation_w_secondary_edges.csv',row.names = F)
}
```
